Fast Sorting on Modern Computers

### 2005-08-14 docs/tech_report/elementary.tex
fixed use of =comparison-based (no hyphen) and fixed up nearly all \n vs \cc usage
fixed usage of '32 byte' without a hyphen

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating an 8KB data
cache with a 32 byte cache line and separate instruction cache.]
{\label{elementary level 1 misses}\includegraphics[scale=\myscale]{plots/on2_-_Level_1_misses.eps}}
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}.]
{\label{elementary branches}\includegraphics[scale=\myscale]{plots/on2_-_Branches.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Cache and Branch prediction simulation results for elementary sorts}
\label{Cache and Branch prediction simulation results for elementary sorts}
\end{figure}
\newpage

>>>>

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating an 8KB data
cache with a 32-byte cache line and separate instruction cache.]
{\label{elementary level 1 misses}\includegraphics[scale=\myscale]{plots/on2_-_Level_1_misses.eps}}
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}.]
{\label{elementary branches}\includegraphics[scale=\myscale]{plots/on2_-_Branches.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Cache and Branch prediction simulation results for elementary sorts}
\label{Cache and Branch prediction simulation results for elementary sorts}
\end{figure}
\newpage

### 2005-08-14 docs/tech_report/report.tex
fixed all references
fixed use of =comparison-based (no hyphen) and fixed up nearly all \n vs \cc usage
finished \cc vs \n search
fixed usage of '32 byte' without a hyphen

%cc is used for
% - unix commans or programs: gcc, sim-bpred etc
\newcommand{\cc}[1]{\texttt{#1}}

%n is used for
% - proper names: Atom, SimpleScalar, Papi
\newcommand{\n}[1]{\textit{#1}}

@@

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating an 8KB data
cache with a 32 byte cache line and separate instruction cache. #4]
{\label{#1 level 1 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_1_misses.eps}}\\
\subfigure[Level 2 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating a 2MB data
and instruction cache and a 32 byte cache line. #5]
{\label{#1 level 2 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_2_misses.eps}}\\
\end{changemargin}
\vspace*{7em}
\caption{Cache Simulation results for #1sort}
\label{Cache Simulation results for #1sort}
\end{figure}
\newpage

@@

\begin{abstract}
Sorting is one of the most important and studied problems in computer
science. Many good algorithms exist which offer various trade-offs in
efficiency, simplicity and memory use. However most of these
algorithms were discovered decades ago at a time when computer
architectures were much simpler than today. Branch prediction and
cache memories are two developments in computer architecture that have
a particularly large impact on the performance of sorting algorithms.

@@

We also developed a number of other improvements to the algorithms,
such as removing the need for a sentinel in heapsort. Overall, we
found that a cache-optimized radix sort was the fastest algorithm for
all large arrays; the absence of comparison branches means that the
algorithm causes almost no branch mispredictions.
\end{abstract}

>>>>

\newenvironment{spacedpars}{%for the abstract
\setlength{\parskip}{1em}
\setlength{\parindent}{0ex}
}

%cc is used for
% - unix commans or programs: gcc, sim-bpred, sim-cache, espresso, li, simple,
%   sim-fast, diff, valgrind, cachegrind, driver
% - code or keywords: less, Item, exch, if, while, unsigned int, int, qsort,
%   goto, do-while, THRESHOLD, partition, malloc, shellsort, improved shellsort
% - file names: main.c
% - #defines: UNIT_MAX
\newcommand{\cc}[1]{\texttt{#1}}

%n is used for
% - proper names: Atom, SimpleScalar, Papi, GnuPlot, ValGrind, glibc
% - technical names: unsigned integer, less-than, greater-than, in-place, out-of
%   place, Abstract Data Type, records, cache hierarchy, latency, hit, miss,
%   compulsory, cold-start, capacity, conflict, cache line, cache block, offset,
%   tag, virtaul addresses, bus address, temporal locality, locality, spatial
%   locality, taken, not taken, flow control branch, comparaitve branch, static,
%   semi-static, dynamic, 1-bit, bimodal, two-level adaptive, heap, children,
%   implicit heap, heapifying, sifting, fix-up, fix-down, sift-up, sift-down,
%   Repeated-adds, Remove-min, d-heap, fanout, natural, straight, unrolled,
%   Memory management unit, tiling, k-way merge, trashing divide and conquer,
%   multi-way partition, comparision based, counting based, increments
% - web addresses: http://www.gnuplot.com, http://valgrind.kde.org
% - sort names, when highlighted: algorithm S, algorithm N
% - branch predictor names: 'Creation, child promotion', etc
\newcommand{\n}[1]{\textit{#1}}

@@

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating an 8KB data
cache with a 32-byte cache line and separate instruction cache. #4]
{\label{#1 level 1 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_1_misses.eps}}\\
\subfigure[Level 2 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating a 2MB data
and instruction cache and a 32-byte cache line. #5]
{\label{#1 level 2 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_2_misses.eps}}\\
\end{changemargin}
\vspace*{7em}
\caption{Cache Simulation results for #1sort}
\label{Cache Simulation results for #1sort}
\end{figure}
\newpage

@@

\begin{spacedpars}
\begin{abstract}
Sorting is one of the most important and studied problems in computer
science. Many good algorithms exist which offer various trade-offs in
efficiency, simplicity and memory use. However most of these
algorithms were discovered decades ago at a time when computer
architectures were much simpler than today. Branch prediction and
cache memories are two developments in computer architecture that have
a particularly large impact on the performance of sorting algorithms.

@@

We also developed a number of other improvements to the algorithms,
such as removing the need for a sentinel in heapsort. Overall, we
found that a cache-optimized radix sort was the fastest algorithm for
all large arrays; the absence of comparison branches means that the
algorithm causes almost no branch mispredictions.
\end{abstract}
\end{spacedpars}
